---
title: "p8105 HW4 Plots"
author: Meghan Bellerose
date: October 27, 2020
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

# Instacart 

The following plots were generated using a random sample of the 2017 Instacart Online Grocery Orders dataset. 

### Setup and data download

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(plotly)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "left"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
data("instacart")

instacart = 
  instacart %>% 
  select(order_id, order_hour_of_day, reordered, user_id, product_id, product_name, department) %>% 
  sample_n(10000)
```

The sample contains `r nrow(instacart)` randomly selected rows and `r ncol(instacart)` columns, with each row representing a single product from an instacart order. Variables include identifiers for the order and product, the day and time of the order, the grocery department, and the product name. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders.

The top 20 bestselling items are listed in the table below:

```{r}
instacart %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 20) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

## Top selling departments

This plot shows the number of items ordered by department.

```{r}
instacart %>% 
  count(department) %>% 
  mutate(
    department = fct_reorder(department, n)) %>%
  plot_ly(
    x = ~department, y = ~n, color = ~department,
    type = "bar", colors = "viridis")
```

Produce, dairy/eggs, and snacks are the top 3 departments from which orders come.

## Boxplot

This plot shows the time of day (hour 0-24) when orderes were placed by department.
```{r}
instacart %>% 
   mutate(
    department = fct_reorder(department, order_hour_of_day)) %>%
  plot_ly(
    x = ~department, y = ~order_hour_of_day, color = ~department,
    type = "box", colors = "viridis")
```

Most orders are placed in the late morning and the afternoon. Most departments see orders throughout the day, with the exception of a few departments, such as alcohol (where items are rarely ordered before 9am).


## Proprtion reordered

This scatterplot shows the proportion of products that are re-ordered. 

```{r}
instacart %>% 
  group_by(product_name) %>% 
  summarize(proportion_reordered = mean(reordered), n=n()) %>% 
  plot_ly(
    x = ~n, y = ~proportion_reordered,
    alpha = .5, type = "scatter", mode = "markers")
```

The plot indicates that, as we would expect, products that are ordered most often are also re-ordered most often, but there is a tapering effect.  